Extended Data

### 2022-03-13 doc/manuscript/Dow_Extended_Data.Rmd
update criteria for CTW selection for leaf phenology
@camerondow35 , I think the better criteria to state here is that we used the group containing the dominant species at each site. Liriodendron accounts for almost 1/2 of productivity at SCBI (and probably more than half of canopy productivity), and red oak is probably >90% at HF.

```{r mean_growth_parameters_table,  eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
mean_growth_parameters <- read.csv("tables_figures/phenology_parameters.csv",  col.names = c("", "ring-porous","diffuse-porous","ring-porous", "diffuse-porous"), stringsAsFactors = FALSE, check.names = FALSE)
kable(mean_growth_parameters, caption = mean_growth_parameters_caption ,  format = "latex", booktabs = TRUE,  escape = F) %>%
add_header_above(c(" ",  "SCBI" = 2, "Harvard Forest"=2 )) %>%
    column_spec(1, width = "4cm") %>%
    pack_rows("Stem Growth", 2, 7, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
    pack_rows("Canopy Foliage (ecosystem level)", 8, 11, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
    pack_rows("Temperature sensitivity* (days/°C)", 12, 20, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
    add_indent(13:16)%>%
    add_indent(18:21)%>%
kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```
$^\ast$Temperature sensitivity refers to the sensitivity of a parameter to the mean maximum temperature ($T_{max}$) during the critical temperature window (Fig. 1). Because leaf phenology measurements were derived from satellite imagery and included both ring- and diffuse-porous trees, the selection of a critical window was less straightforward. We used the critical window which had the greatest correlation with leaf greenup timing as the critical window for leaf phenology calculations. This resulted in the selection of the diffuse-porous critical window at SCBI and the ring-porous critical window at HF, corresponding to the grouping of the species with greatest woody productivity at each site (diffuse-porous *Liriodendron tulipifera* at SCBI, ring-porious *Quercus rubra* at HF).

>>>>

```{r mean_growth_parameters_table,  eval=TRUE, echo=FALSE, warning=FALSE}
library(knitr)
library(kableExtra)
mean_growth_parameters <- read.csv("tables_figures/phenology_parameters.csv",  col.names = c("", "ring-porous","diffuse-porous","ring-porous", "diffuse-porous"), stringsAsFactors = FALSE, check.names = FALSE)
kable(mean_growth_parameters, caption = mean_growth_parameters_caption ,  format = "latex", booktabs = TRUE,  escape = F) %>%
add_header_above(c(" ",  "SCBI" = 2, "Harvard Forest"=2 )) %>%
    column_spec(1, width = "4cm") %>%
    pack_rows("Stem Growth", 2, 7, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
    pack_rows("Canopy Foliage (ecosystem level)", 8, 11, latex_gap_space = "1em", colnum = 4, hline_before = FALSE) %>%
    pack_rows("Temperature sensitivity* (days/°C)", 12, 20, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
    add_indent(13:16)%>%
    add_indent(18:21)%>%
kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```
$^\ast$Temperature sensitivity refers to the sensitivity of a parameter to the mean maximum temperature ($T_{max}$) during the critical temperature window (CTW, Fig. 1). Because leaf phenology measurements were derived from satellite imagery and included both ring- and diffuse-porous trees, the selection of a critical window was less straightforward. We used the CTW of the porosity group containing the dominant canopy species at each site: *Liriodendron tulipifera* (diffuse porous) at SCBI and *Quercus rubra* (ring porous) at Harvard Forest.

### 2022-03-13 doc/manuscript/Dow_manuscript_Nature.Rmd
update criteria for CTW selection for leaf phenology
@camerondow35 , I think the better criteria to state here is that we used the group containing the dominant species at each site. Liriodendron accounts for almost 1/2 of productivity at SCBI (and probably more than half of canopy productivity), and red oak is probably >90% at HF.

![**Figure 2 | Responses of foliage phenology (a,b) and stem growth timing (c,d) to spring temperatures at the Smithsonian Conservation Biology Institute (a,c) and Harvard Forest (b,d).** Panels (a-b) show ecosystem-level canopy foliage phenology from 2001-2018, obtained from the MODIS Global Vegetation Phenology product (MCD12Q2.006) for a single pixel at the center of each ForestGEO plot, where G = Greenup, M=Mid-greenup, P=peak, and S=Senescence (i.e., beginning of green-down). Panels (c-d) show the dates at which stem growth milestones were achieved, on average, for sampled populations of ring porous and diffuse porous trees at SCBI (2011-2020) and Harvard Forest (1999-2003). Mean $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ were estimated using the Bayesian model visualized, with confidence intervals, in `r EDF_pheno_Tmax_climwin`. Mean maximum temperature ($T_{max}$) was calculated for each xylem architecture/site combination over the respective critical temperature window (CTW), then turned into a ratio and assigned a color on a gradient where the coldest year in the sample is blue and the warmest is red. Leaf phenology years are colored according to the CTW $T_{max}$ with the best correlation with timing of greenup (diffuse porous window at SCBI and ring porous window at Harvard Forest).](tables_figures/DOYtiming_allyears.png){ width=25% }

@@

Whereas the length of time between canopy greenup and senescence (*i.e.,* the day when greenness dropped below 90% of its peak) increased with CTW $T_{max}$ (Fig. `r DOY_timing`a-b)<!--specify which species group used?-->, there was no consistent lengthening of $L_{PGS}$ (Fig. `r schematic_summary`, `r EDF_pheno_all`).<!--Note to Cam: double check this-->

>>>>

![**Figure 2 | Responses of foliage phenology (a,b) and stem growth timing (c,d) to spring temperatures at the Smithsonian Conservation Biology Institute (a,c) and Harvard Forest (b,d).** Panels (a-b) show ecosystem-level canopy foliage phenology from 2001-2018, obtained from the MODIS Global Vegetation Phenology product (MCD12Q2.006) for a single pixel at the center of each ForestGEO plot, where G = Greenup, M=Mid-greenup, P=peak, and S=Senescence (i.e., beginning of green-down). Panels (c-d) show the dates at which stem growth milestones were achieved, on average, for sampled populations of ring porous and diffuse porous trees at SCBI (2011-2020) and Harvard Forest (1999-2003). Mean $DOY_{25}$, $DOY_{50}$, and $DOY_{75}$ were estimated using the Bayesian model visualized, with confidence intervals, in `r EDF_pheno_Tmax_climwin`. Mean maximum temperature ($T_{max}$) was calculated for each xylem architecture/site combination over the respective critical temperature window (CTW), then turned into a ratio and assigned a color on a gradient where the coldest year in the sample is blue and the warmest is red. Leaf phenology years are colored according to the CTW $T_{max}$ of the porosity group containing the dominant canopy species at each site (diffuse porous at SCBI, ring porous at Harvard Forest).](tables_figures/DOYtiming_allyears.png){ width=25% }

@@

Whereas the length of time between canopy greenup and senescence (*i.e.,* the day when greenness dropped below 90% of its peak) increased with the CTW $T_{max}$ of the porosity group containing the dominant canopy species at each site (diffuse porous at SCBI, ring porous at Harvard Forest; Fig. `r DOY_timing`a-b), there was no consistent lengthening of $L_{PGS}$ (Fig. `r schematic_summary`, `r EDF_pheno_all`).<!--Note to Cam: double check this-->
